---
title: "Run Models"
author: "Eliot Boulaire"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true 
  pdf_document:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(root.dir = rstudioapi::getActiveProject(), echo = TRUE)
rm(list = ls())

pkginstall <- function(packages) {
  for (pkg in packages) {
    if (!require(pkg, character.only = TRUE)) {
      message(paste("Installation du package manquant:", pkg))
      install.packages(pkg, dependencies = TRUE)
      library(pkg, character.only = TRUE)
      message(paste(pkg, "chargé avec succès"))
    } else {
      message(paste("Le package", pkg, "est déjà installé."))
      library(pkg, character.only = TRUE)
      message(paste(pkg, "chargé avec succès"))
    }
  }
}

packages <- c("nimble", "coda")
pkginstall(packages)

HOME <- rstudioapi::getActiveProject()
```

# Introduction

## Load data

```{r datas}
data <- readRDS(paste0(HOME, "/Project/data/realdata/data.rds"))
const <- readRDS(paste0(HOME, "/Project/data/realdata/const.rds"))
```

## Load initials values

```{r inits}
inits_3chains <- readRDS(paste0(HOME,"/Project/data/realdata/inits_3chains.rds"))
inits_1chain <- readRDS(paste0(HOME, "/Project/data/realdata/inits_1chain.rds"))
```

## Define monitored values

```{r monitor}
monitor <- c(
  "Mu_L3", "Mu_L41", "Mu_L42", "Mu_L5", "Mu_L8", "Mu_L6", "Mu_L9",
  "Sd_L3", "Sd_L41", "Sd_L42", "Sd_L5", "Sd_L8", "Sd_L6", "Sd_L9",
  "Mu_N3", "Mu_N6", "Mu_N9",
  "Psex_SR3", "Psex_SR6", "Psex_SR9",
  "Alpha3", "Beta3", "Alpha4", "Beta4"
)
```

## Load the model

```{r model}
source(paste0(HOME, "/Project/models/SALMSIZE-MARSIM.R"))
```

## Load the NimbleFunctions used inside the model

```{r functions}
source(paste0(HOME, "/Project/functions/nf_omega.R"))
source(paste0(HOME, "/Project/functions/nf_midbins.R"))
source(paste0(HOME, "/Project/functions/nf_prop.R"))
```

## Define MCMC settings

```{r MCMC}
n_chains = 3
n_thin = 30
n_keep = 1000
n_burnin = 20000
n_iter = n_keep * n_thin + n_burnin
```

# Run Bayesian Models (with NIMBLE)

Steps to perform MCMC estimations on hierarchical bayesian model using NIMBLE system

## Step 1 : Build nimble model

```{r buildmodel}
model_nimble <- nimbleModel(
  code = model_code,
  name = 'model_nimble',
  constants = const,
  data = data,
  inits = inits_1chain
)
# model_nimble$initializeInfo()
# model_nimble$calculate()
```

## Step 2 : Compile the model

```{r compilemodel}
compiled_model <- compileNimble(model_nimble)
```

## Step 3 : Create a MCMC sampler for the model

```{r MCMCsampler}
# Here we use the RAW sampler (base sampler in nimble)
model_conf <- configureMCMC(
  model = model_nimble,
  thin = 1,
  monitors = monitor,
  inits = inits_1chain,
  enableWAIC = TRUE
)

model_MCMC <- buildMCMC(model_conf)
# model_MCMC$run(niter = 0)
```

## Step 4 : Compile the existing model with the associated MCMC sampler

```{r compileAll}
compiled_MCMC <- compileNimble(model_MCMC, project = model_nimble)
```

## Step 5 : Run MCMC

```{r runMCMC}
set.seed(123)

MCMC_samples <- runMCMC(
  mcmc = compiled_MCMC,
  niter = n_iter,
  nburnin = n_burnin,
  nchains = n_chains,
  thin = n_thin,
  inits = inits_3chains,
  progressBar = TRUE,
  samples = TRUE,
  samplesAsCodaMCMC = TRUE,
  summary = FALSE,
  WAIC = TRUE # FALSE you don't want WAIC
)
```

## Step 6 : Save MCMC

```{r saveMCMC}
is(MCMC_samples$samples)
varnames(MCMC_samples$samples)

saveRDS(MCMC_samples, file = paste0(HOME, "/Project/saves/MCMC_base.rds")) # Change the name of the file as you want
```
